---
title: "Five pool model fitting: Fixed Initial Input"
author: "S. Y. Luk (sheronyl@mit.edu) and K. Todd-Brown (ktoddbrown@gmail.com)"
date: "8/7/2019"
output:
  pdf_document: default
  html_document: default
---


# Introduction
In the two pool model, we estimated turnover times of fine SOC in the marsh platform using a depth-for-time substitution where down-core changes in carbon concentration were modeled against geochronology. The two-pool model was developed from the persistent low and high temperature thermogram peaks from ramped pyrolysis oxidation of Plum Island Ecosystem - Long Term Ecological Reserve domain marsh soils. Turnover of the first pool is described as first-order kinetics dictated by the carbon concentration of the pool at a given time (obtained by age models), linear input (U) during the time within the root zone (troot) and turnover time. Subsequent pools are a linear series where a fraction (r) of decomposed material from the previous pool is transferred to the next pool. 

```{r setup}
library(tidyverse)
library(deSolve) #lsoda
library(tidyr)
library(dplyr)
library(kableExtra)
```

#Data Description
Here, we load the two pool data from ramped pyrolysis oxidation. Due to large age estimates for each soil horizon we included the mean, min, max of each soil horizon in the dataset to be estimated by the model. Time is defined as the soil age horizon estimate. The first and second pool in mg C per g dry sediment is defined in C1 and C2, respectively. A unique identifier for each soil horizon (ID) was provided, which each have a min (REP=1), mean (REP=2), max(REP=3) age estimate.
```{r readData}
#Load data file
data.df<- read_csv(file = file.path('data', "TwoPoolDataFinal.csv"), 
                   skip = 1, #skip the headers
                   col_names = c('time','C1','C2','ID','REP')) %>% #give the columes better names
  arrange(time)
```

Turnover of the first pool is described as first-order kinetics dictated by the carbon concentration of the pool at a given time (t, obtained by age models), linear input (U) during the time within the root zone (troot) and turnover time (tau). Subsequent pools are a linear series where a fraction (r) of decomposed material from the previous pool is transferred to the next pool. 

$U_i(t)=U_{initial,i}-U_{initial,i} \frac{t}{t_{root}} \text{   if } t<t_{root}$

$\frac{dx_1 (t)}{dt} = U_1 (t)-\frac{1}{\tau_1}x_1 (t)$

```{r modelFunc_with_inputs}
NPoolModel_modInput.fn <- function(t, y, parms, ...){
  #for linear fit I_max+(-I_max/root_time)*t
  dy <- ifelse(t < parms$root_time, 
               parms$I_max - parms$I_max / parms$root_time * t,
               0) + #if we are before the root zone input 'time period' add roots
    parms$decay %*% y[1:nrow(parms$decay)] #otherwise decay
  return(list(c(dy)))
}
```

# Parameter runs and selection

```{r fitFirstPool}
if(!file.exists(file.path('outputs', "TwoPoolFirstSecondPool.Rdata"))){
  #Generate combination of parameters for model to run through: decay rate (k), starting_C (x0), transfer (r), input (I0), over input time within rooting zone defined as 85 years 
  firstPool <- expand.grid(list(decay_rates = 1/seq(0.5, 10000, by=10),
                                starting_C = seq(75, 140, by=1),
                                r=0,
                                input=seq(0, 3, by=0.5),
                                inputtime=85 #years for 25cm root depth
  )) %>%
    group_by_all() %>%
    do((function(xx){
      lsoda(y=c(C=xx$starting_C), 
            times = data.df$time, 
            func = NPoolModel_modInput.fn,
            parms = list(I_max=xx$input, root_time=xx$inputtime, decay = matrix(-xx$decay_rates), transfer = xx$r)) %>%
        as.data.frame %>% 
        mutate(C_data = data.df$C1, ID=data.df$ID, REP=data.df$REP)%>%
        ungroup()%>%
        dplyr::select(C,C_data,ID,REP) %>%
        do((function(x){
          data.frame(RMSE=sqrt(sum((x$C-x$C_data)^2)/length(x$C_data)),
                     broom::tidy(lm(x$C_data ~ x$C)))
        })(.))
    })(.)) %>%
    #Filter acceptable model fits based on linear regression of observed and modeled output. Good fit is whether intercept+/-SE overlap 0 and slope+/- overlap 1.
    filter(estimate[term == '(Intercept)'] - std.error[term == '(Intercept)'] < 0 &
             estimate[term == '(Intercept)'] + std.error[term == '(Intercept)'] > 0) %>%
    filter(estimate[term == 'x$C'] - std.error[term == 'x$C'] < 1 &
             estimate[term == 'x$C'] + std.error[term == 'x$C'] > 1 ) 
  
  set.seed(123)
  #Filter based on lowest quartile RMSE
  firstPool.parm <- firstPool %>% 
    filter(RMSE<summary(firstPool$RMSE)[[2]]) %>%
    dplyr::select(r, decay_rates, starting_C, input,inputtime,RMSE) %>% 
    unique() %>% ungroup() %>%
    mutate(C1_index = 1:nrow(.)) %>%
    dplyr::rename('C1_r' = 'r','C1_decay' = 'decay_rates', 'C1_0' = 'starting_C', 'C1_in' = 'input','C1_inputtime'='inputtime','C1_RMSE'='RMSE') %>%
    #Random sampling of 50 inclusive of the minimum RMSE model run
    sample_n(50)
  
  #Add RMSE minimum 
  firstPool.parm<-rbind(firstPool.parm[firstPool.parm$C1_RMSE == min(firstPool.parm$C1_RMSE),],firstPool.parm) %>%
    unique()%>%
    mutate(C1_index = 1:nrow(.))
  
  
  secondPool <- expand.grid(list(decay_rates = 1/seq(0.5, 10000, by=50),
                                 starting_C = seq(20, 50, by=1),
                                 input=seq(0, 3, by=0.5), 
                                 inputtime=85, #years for 25cm root depth
                                 r = seq(0,1, by=0.5),
                                 C1_index = firstPool.parm$C1_index)) %>% 
    group_by_all() %>%
    do((function(xx){
      lsoda(y=c(C1 = firstPool.parm$C1_0[xx$C1_index], C=xx$starting_C), 
            times = data.df$time, 
            func = NPoolModel_modInput.fn,
            parms = list(I_max=c(firstPool.parm$C1_in[xx$C1_index], xx$input), 
                         root_time=c(firstPool.parm$C1_inputtime[xx$C1_index], xx$inputtime),  
                         decay = matrix(c(-firstPool.parm$C1_decay[xx$C1_index], 0,
                                          firstPool.parm$C1_decay[xx$C1_index]*xx$r,-xx$decay_rates), nrow = 2, byrow = TRUE))) %>%
        as.data.frame %>% 
        mutate(C_data = data.df$C2, ID=data.df$ID, REP=data.df$REP) %>%
        ungroup()%>%
        dplyr::select(C,C_data,ID,REP)%>%
        do((function(x){
          data.frame(RMSE=sqrt(sum((x$C-x$C_data)^2)/length(x$C_data)),
                     broom::tidy(lm(x$C_data ~ x$C)))
        })(.))
    })(.)) %>% 
    filter(estimate[term == '(Intercept)'] - std.error[term == '(Intercept)'] < 0 &
             estimate[term == '(Intercept)'] + std.error[term == '(Intercept)'] > 0) %>%
    filter(estimate[term == 'x$C'] - std.error[term == 'x$C'] < 1 &
             estimate[term == 'x$C'] + std.error[term == 'x$C'] > 1 ) 
  
  set.seed(123)
  secondPool.parm <- secondPool %>% 
    filter(RMSE<summary(secondPool$RMSE)[[2]]) %>% 
    ungroup() %>%
    dplyr::select(r,decay_rates,starting_C,input,inputtime,RMSE,C1_index) %>% 
    unique() %>% 
    rename('C2_r'='r','C2_decay' = 'decay_rates', 'C2_0' = 'starting_C', 'C2_in' = 'input', 'C2_inputtime' = 'inputtime','C2_RMSE'='RMSE') %>%
    left_join(firstPool.parm) %>%
    sample_n(50)
  
  secondPool.parm<-rbind(secondPool.parm[secondPool.parm$C2_RMSE == min(secondPool.parm$C2_RMSE),],secondPool.parm) %>%
    unique()%>%
    mutate(C2_index = 1:nrow(.))
  
  save.image(file=file.path('outputs', "TwoPoolFirstSecondPool.Rdata"))
}else{
  load(file.path('outputs', "TwoPoolFirstSecondPool.Rdata"))
}

```

# Best fit simulation reruns

```{r modelDataPlot}


#run simulation of all good model fits
simulations <- secondPool.parm %>% 
  full_join(firstPool.parm) %>% ungroup() %>% 
  mutate(index = 1:nrow(.)) %>%
  group_by_all() %>%
  do((function(xx){
    lsoda(y=c(C1 = firstPool.parm$C1_0[xx$C1_index], C2=xx$C2_0), 
          times = data.df$time, 
          func = NPoolModel_modInput.fn,
          parms = list(I_max=c(firstPool.parm$C1_in[xx$C1_index], xx$C2_in), 
                       root_time=c(firstPool.parm$C1_inputtime[xx$C1_index], xx$C2_inputtime), 
                       decay = matrix(c(-firstPool.parm$C1_decay[xx$C1_index], 0,
                                        firstPool.parm$C1_decay[xx$C1_index]*xx$C2_r,-xx$C2_decay), nrow = 2, byrow = TRUE))) %>%
      as.data.frame
  })(.))

#select best fit model based on minimizing sum of RMSE
secondPool.parm$sumRMSE<-secondPool.parm$C1_RMSE+secondPool.parm$C2_RMSE
bestFitfinal<-secondPool.parm[secondPool.parm$sumRMSE==min(secondPool.parm$sumRMSE),]

if(!file.exists(file.path('outputs', 'TwoPoolBestFit.csv'))){
  bestfitsimulations <- bestFitfinal %>% 
    left_join(firstPool.parm) %>% ungroup() %>% 
    mutate(index = 1:nrow(.)) %>%
    group_by_all() %>%
    do((function(xx){
      lsoda(y=c(C1 = xx$C1_0, C2=xx$C2_0), 
            times = data.df$time, 
            func = NPoolModel_modInput.fn,
            parms = list(I_max=c(xx$C1_in, xx$C2_in), 
                         root_time=c(xx$C1_inputtime, xx$C2_inputtime), 
                         decay = matrix(c(-xx$C1_decay, 0,
                                          xx$C1_decay*xx$C2_r,-xx$C2_decay), nrow = 2, byrow = TRUE))) %>%
        as.data.frame
    })(.))
  
  write.csv(bestfitsimulations, file="TwoPoolBestFit.csv") #write to create the figure in matlab
}else{
  bestfitsimulations <- read_csv(file.path('outputs', 'TwoPoolBestFit.csv'))
}
```

# Results

```{r plot}
ggplot(data.df %>% pivot_longer(cols = c('C1','C2'), names_to = 'pool'),
       aes(x=time, y=value))+
  geom_line(data = simulations %>% ungroup() %>% dplyr::select(index, time,C1,C2) %>%
              pivot_longer(cols = c('C1','C2'), names_to = 'pool'),
            aes(x=time, y=value, color = pool, group=paste(index, pool)), alpha = 0.1) +
  
     geom_line(data = bestfitsimulations %>% ungroup() %>% dplyr::select(index, time, C1, C2) %>%
              pivot_longer(cols = c('C1','C2'), names_to = 'pool'),
            aes(x=time, y=value, color = 'Best Fit', group=paste(index, pool)), alpha = 1, lwd=2) +
  
  geom_point() +
  xlab('Time') + ylab("CO2 Evolved (mg C"  ~ g^{-1} ~ "dry sediment)") +
  theme_bw() +
  facet_wrap(~pool)
```

# Summary Table of results for best fit, min, and max estimates for turnover time (tau), root inputs (U), initial pool size (x0), and fraction transferred from pool 1 (r).
```{r summary table}
temp1 <- firstPool.parm %>%
  mutate(C1_tau = 1/C1_decay) %>%
  pivot_longer(cols = c('C1_tau', 'C1_0', 'C1_in')) %>%
  group_by(name) %>%
  dplyr::summarise(max_value = max(value), 
            min_value = min(value),
            bestRMSE = value[unique(bestfitsimulations$C1_index) == C1_index]) %>%
  ungroup() %>%
  mutate(name = recode(name,
                       C1_tau = 'Pool 1 turnover time [yr]',
                       C1_in = 'Pool 1 root zone input [mg-C per g-dry_soil]',
                       C1_0 = 'Inital pool 1 size [mg-C per g-dry_soil]'))

temp2 <- secondPool.parm %>%
  mutate(C2_tau = 1/C2_decay) %>%
  pivot_longer(cols = c('C2_r', 'C2_tau', 'C2_0', 'C2_in')) %>%
  group_by(name) %>%
  dplyr::summarise(max_value = max(value), 
            min_value = min(value),
            bestRMSE = value[unique(bestfitsimulations$C2_index) == C2_index])%>%
  ungroup() %>%
  mutate(name = recode(name,
                       C2_r = 'Transfer from pool 1 to pool 2 [fraction]',
                       C2_tau = 'Pool 2 turnover time [yr]',
                       C2_in = 'Pool 2 root zone input [mg-C per g-dry_soil]',
                       C2_0 = 'Inital pool 2 size [mg-C per g-dry_soil]'))

ranges <- bind_rows(temp1, temp2)

ranges %>%
  kbl() %>%
  kable_classic_2(full_width = F)

write.csv(ranges, file=file.path('outputs', "TwoPoolBestFitRanges.csv"))
```